Social Participation Ontology: community documentation, enhancements and use examples

### 2015-09-21 plosOne/plos.tex
revisadas primeiras 2 secoes

\section{\vcps\ original documentation}
From April to December, 2012, \vcps\ was first conceived. In the online process, as registered by \corais\ platform, 66 users interacted, 6 of them were the most active~\cite{metodologia}. Various materials were produced both as activity traces and as reference media. This section is dedicated to these materials.

\subsection{Reference textual documents}\label{refDocs}
The main documents are:
\begin{itemize}
    \item ``Commented methodology''~\cite{metodologia}: this document describes the public process of \vcps\ conception. It is composed by brief inspections of forum topics, pointing both pertinent characteristics of the online collective process and ontological observations (about classes and properties). Considerations are made about tightening relations with the Open Government Partnership (\ogp), an international initiative to foster transparency and open practices in governments worldwide~\cite{OGP}, and the Brazilian formal action plan, as means to achieve ontology usage. 
     There is also a proposal of a systematic study of electronic government initiatives, so that the \vcps\ might be better contextualized. This document ends by proposing an agenda of meetings with academics, entrepreneurs and government parties.
 \item ``Conceptual modeling, version 0.1 (in natural language)''~\cite{conceptualMod}: this document is a description, in common English, of the \vcps. The introduction is mainly a collage of the document above~\cite{metodologia}. Both the itemized description of the ontology, and the considerations for its usage, are of great value as reference. Figure~\ref{fig:v1} is heavily influenced by a diagram related to this document and further described in Section~\ref{sec:im}.
\end{itemize}

@@

\begin{itemize}
    \item Various proposals for the \vcps\ logo, some of which are in Figure~\ref{logo}.
    \item Figure~\ref{fig:v1} shows an English and completed version of the original diagram of \ops\ in the document~\cite{conceptualMod}. For completeness of exposition, the original diagram is Supporting Information Figure~\ref{fig:diaorig}.
    \item A diagram for general public consultations. Given the details and the pertinence of public consultations for \ops, the digram is exposed in Supporting Information Figure~\ref{fig:consult}.
\end{itemize}

\begin{figure}[!h]
    \centering
%    \includegraphics[width=0.4\textwidth]{figs/logoUnificadoDoPDF}
    \caption{Some of the various logos for of the \vcps. (a) is a colored text logo proposal; (b) is a figurative logo; (c) is mixture of both ideas. It can be seen that these logos were conceived for the ontology when it was called a \emph{vocabulary}. Community documents reflect this nomenclature, which is being changed with this article and subsequent work with \ops\ advent.}
    \label{logo}
\end{figure}

@@

\subsection{Blog Posts}
The \vcps\ blog
aggregates both important discussions and documents in no more than twenty posts to date.
All \owl\ code, final documents, public consultations, mental map and images are posted in the blog~\cite{coraisBlog}. 
The first post if from July 24, 2012. Last post about \vcps\ is from May 7, 2013. Most blog posts are from the first day (almost half of them). They received more than
twenty commentaries. Two ``out-of-season'' blog posts, one from August 9, 2012 and another from October  22, 2012, separate first day posts from last posts. Both have about ten commentaries.
Last blog posts occurred as a few days burst and a final message, a month after.

@@

\section{\ops: the Ontology of Social Participation}\label{exp}
%A próxima secao já é a conclusão. Alem dessa sessão ter ficado grande, voce precisa uma sessão onde demonstra que essa ontologia é util para alguma coisa

\begin{figure*}
    \centering
%    \includegraphics[width=0.9\textwidth]{figs/opsBFO___}
    \caption{Diagram representation of the Ontology of Social Participation (\ops). Arrows with white heads indicate ``is a'' relations (subclass points to superclass). Arrows with black heads indicate property relations from subject to object . All property relation yield existential restrictions, with the exception of the ``has characteristic'' property, that does not yield restriction. Upper ontologies \bfo\ and \foaf\ are asserted through color (\bfo) and dashed boxes (\foaf).}
    \label{fig:v1}
\end{figure*}

>>>>

\section{VCPS original documentation}
From April to December, 2012, \vcps\ was first conceived. In the online process, as registered by \corais\ platform, 66 users interacted, 6 of them were the most active~\cite{metodologia}. Various materials were produced both as activity traces and as reference media. This section is dedicated to these materials.

\subsection{Reference textual documents}\label{refDocs}
The main documents are:
\begin{itemize}
    \item ``Commented methodology''~\cite{metodologia}: this document describes the public process of \vcps\ conception. It is composed by brief inspections of forum topics, pointing both pertinent characteristics of the online collective process and ontological observations (about classes and properties). Considerations are made about tightening relations with the Open Government Partnership (\ogp), an international initiative to foster transparency and open practices in governments worldwide~\cite{OGP}, and the Brazilian formal action plan, as means to achieve ontology usage. 
     There is also a proposal of a systematic study of electronic government initiatives, so that the \vcps\ might be better contextualized. This document ends by proposing an agenda of meetings with academics, entrepreneurs and government parties.
 \item ``Conceptual modeling, version 0.1 (in natural language)''~\cite{conceptualMod}: this document is a description, in common English, of the \vcps. The introduction is mainly a collage of the document above~\cite{metodologia}. Both the itemized description of the ontology, and the considerations for its usage, are of great value as references. Figure~\ref{fig:v1} is heavily influenced by a diagram related to this document and further described in Section~\ref{sec:im}.
\end{itemize}

@@

\begin{itemize}
    \item Various proposals for the \vcps\ logo, some of which are in Figure~\ref{logo}.
    \item Figure~\ref{fig:v1} shows an English and completed version of the original diagram of \ops\ in the document~\cite{conceptualMod}. For completeness of exposition, the original diagram is in Supporting Information Figure~\ref{fig:diaorig}.
    \item A diagram for general public consultations. Given the details and the pertinence of public consultations for \ops, the digram is exposed in Supporting Information Figure~\ref{fig:consult}.
\end{itemize}

\begin{figure}[!h]
    \centering
%    \includegraphics[width=0.4\textwidth]{figs/logoUnificadoDoPDF}
    \caption{Some of the various logos for of the \vcps. (a) is a colored text logo proposal; (b) is a figurative logo; (c) is mixture of both ideas. It can be seen that these logos were conceived for the ontology when it was called a \emph{vocabulary}. Community documents reflect this nomenclature, which changed with first versions of this article, in the final months of 2013, and subsequent work.}
    \label{logo}
\end{figure}

@@

\subsection{Blog Posts}
The \vcps\ blog
aggregates both important discussions and documents in no more than twenty posts to date.
All \owl\ code, final documents, public consultations, mental map and images are posted in the blog~\cite{coraisBlog}. 
The first post is from July 24, 2012. Last post about \vcps\ is from May 7, 2013. Most blog posts are from the first day (almost half of them). They received more than
twenty commentaries. Two ``out-of-season'' blog posts, one from August 9, 2012 and another from October  22, 2012, separate first day posts from last posts. Both have about ten commentaries.
Last blog posts occurred as a few days burst and a final message, a month after.

@@

\section{OPS: the Ontology of Social Participation}\label{exp}
%A próxima secao já é a conclusão. Alem dessa sessão ter ficado grande, voce precisa uma sessão onde demonstra que essa ontologia é util para alguma coisa

\begin{figure*}
    \centering
%    \includegraphics[width=0.9\textwidth]{figs/opsBFO___}
    \caption{Diagram representation of the Ontology of Social Participation (\ops). Arrows with white heads indicate ``is a'' relations (subclass points to superclass). Arrows with black heads indicate property relations from subject to object . All property relations yield existential restrictions, with the exception of the ``has characteristic'' property, that does not yield restriction. Upper ontologies \bfo\ and \foaf\ are asserted through color (\bfo) and dashed boxes (\foaf).}
    \label{fig:v1}
\end{figure*}
